Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-04-17 closed_loop_planning.tex
[MGa] draft for closed loop

Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to stabilize the deviations along the nominal trajectory. To compute the state-feedback matrices $\hbK_k$, we linearize the system around $(\hbmu_k, \hbu_k)$, such that $\bbmu_{k+1} \approx \hbA_{k}\bbmu_k + \hbB_k \bbu_k$, where $\hbA_k = f_{,\bmu}(\hbmu_k,\hbu_k)$ and $\hbB_k = f_{,\bu}(\hbmu_k,\hbu_k)$, and define the quadratic regulator (tracking) cost function
\begin{align}
J = \sum_{k=0}^{N-1} \big( \bbmu_k^{T} \bW_k \bbmu_k + \bbu_k^T \bR_k \bbu_k \big) + \bbmu_N^T \bW_N \bbmu_N,
\end{align}
where $\bW_N=\bW_N^T \succ 0$, $\bW_k=\bW_k^T \succeq 0$, $\bR_k=\bR_k^T \succ 0$. It can be shown that the optimal cost-to-go is given by $J_k^*(\bbmu_k) = \bbmu_k^T \bS_k \bbmu_k$, with $\bS_k = \bS_k^T \succ 0$, where $\bS_k$ can be computed through the (backward) Riccati recursion for $k=N-1,\ldots, 0$
\begin{align}
\bS_N &= \bW_N,\\ 
\bS_{k} &= \bW_k + \hbA_k^T \bS_{k+1} \hbA_k - \hbA_k^T \bS_{k+1}\hbB_k \hbK_k,\quad k={N-1, \ldots, 0}
\end{align}
where  $\hbK_k =\big( \bR_k+\hbB_k^T \bS_{k+1} \hbB_k \big)^{-1} \hbB_k^T \bS_{k+1} \hbA_k$. The optimal (stabilizing) feedback policy is given by $\bbu_k=-\hbK_k \bbmu_k$. We assume that the gain sequence $\{\hbK_k\}$, $k=0,\ldots, N$ represents the control strategy that an expert driver would employ to optimally stabilize the system along the nominal trajectory.

@@

>>>>

Given the nominal trajectory $(\hbmu_k, \hbu_k)$ over $k=0,\ldots, N$, and introduced the deviation variables $\bbmu_k=\bmu_k - \hbmu_k$ and $\bbu = \bu_k - \hbu_k$, we design a linear state-feedback law $\bbu_k= -\hbK_k \bbmu_k$, or equivalently $\bu_k = \hbu_k - \hbK_k (\bmu_k - \hbmu_k)$, to stabilize the deviations along the nominal trajectory. To compute the state-feedback matrices $\hbK_k$, we linearize the system around $(\hbmu_k, \hbu_k)$, such that $\bbmu_{k+1} \approx \hbA_{k}\bbmu_k + \hbB_k \bbu_k$, where $\hbA_k = f_{,\bmu}(\hbmu_k,\hbu_k)$ and $\hbB_k = f_{,\bu}(\hbmu_k,\hbu_k)$, and define the quadratic regulator (tracking) cost function
\begin{align}
J = \bbmu_N^T \bW_N \bbmu_N + \sum_{k=0}^{N-1} \big( \bbmu_k^{T} \bW_k \bbmu_k + \bbu_k^T \bR_k \bbu_k \big),
\end{align}
where $\bW_N=\bW_N^T \succ 0$, $\bW_k=\bW_k^T \succeq 0$, $\bR_k=\bR_k^T \succ 0$. It can be shown that the optimal cost-to-go is given by $J_k^*(\bbmu_k) = \bbmu_k^T \bS_k \bbmu_k$, with $\bS_k = \bS_k^T \succ 0$, where $\bS_k$ can be computed through the (backward) Riccati recursion for $k=N-1,\ldots, 0$
\begin{align}
\bS_N &= \bW_N,\\ 
\bS_{k} &= \bW_k + \hbA_k^T \bS_{k+1} \hbA_k - \hbA_k^T \bS_{k+1}\hbB_k \hbK_k,\quad k={N-1, \ldots, 0}
\end{align}
where  $\hbK_k =\big( \bR_k+\hbB_k^T \bS_{k+1} \hbB_k \big)^{-1} \hbB_k^T \bS_{k+1} \hbA_k$. The optimal (stabilizing) feedback policy is given by $\bbu_k=-\hbK_k \bbmu_k$. We assume that the gain sequence $\{\hbK_k\}$, $k=0,\ldots, N$ represents the control strategy that an expert driver would employ to optimally stabilize the system along the nominal trajectory.

@@

%\begin{align}
%    \underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}} \quad & J_k(\bmu_k,\bxi_k, \bu_k) \label{eq:ROCPcost}\\
%       \text{s.t.}\notag\\
%          & f_0(x) - t \le 0 \quad \phantom{i=1, \dots, m}\\
%          & f_i(x) \le 0, \quad i=1, \dots, m \quad i=1, \dots, m\\
%          & h_i(x) = 0,   \quad i=1, \dots, p,
%\end{align}

\begin{alignat}{3}
\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}}
    & \quad J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:ROCPcost} \\
\text{s.t.}\notag\\
\bzero      & = \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
\quad k = 1,\ldots, N \label{eq:ROCPdyn} \\
\bmu_0      & = \; \bar{\bmu}_0 \label{eq:ROCPdynIC} \\
\bzero      & = \; \bPsiPCL_k(\bmu_k,\bxi_k, \bu_k, \tbP_{k-1},\tbP_k,\bSi_k,\bz_k, \bK_k),
\quad k = 1,\ldots, N \label{eq:ROCPdP} \\
\tbP_0       & = \; \bar{\bP}_0 \succeq 0 \label{eq:ROCPdPIC} \\
\bzero      & = \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k), \quad k = 0,\ldots, N \label{eq:ROCPpath} \\
0    & \geq  h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \tbP_k, \bz_k), \quad k = 1,\ldots, N;\; \quad i \in \calI \label{eq:ROCPconstraints}\\
-\de \bK   & \leq \bK_k - \hbK_k \leq  \de \bK \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
\end{alignat} 
where~\eqref{eq:ROCPdP} denotes the discretized form for time evolution of the covariance matrix under a stabilizing controller of the form $\bbu_k=-\bK_k \bbmu_k$. As apposed to~\eqref{eq:OCPdP}, $\tbP$ would evolve, in continuous time, according to $\dot{\tbP}(t) = (\bA(t)-\bB(t) \bK(t)) \tbP(t) + \tbP(t) (\bA(t)-\bB(t) \bK(t))^T + \bQ(t)$. To be continued...
